---
author: "Laura Botzet"
date: "9/15/2016"
output:
  html_document:
    lib_dir: library
    self_contained: no
  pdf_document: default
---


## Library
```{r Library}
library(formr)
library(devtools)
library(haven)
library(tidyr)
library(ggplot2)
library(stringr)
library(GPArotation)
library(dplyr)
```

## Helper
```{r helper}
### Function to calculate the birthdate out of all available informations for one individual
all_available_info_birth_date = function(byear, bmonth, bday = NULL) {
  if(!is.null(bday)) {
    bday = paste0("-", bday)
  } else {
    bday = ""
  }
  ifelse(is.na(byear), NA,
         paste0(byear, "-", bmonth, bday))
  # can yield 2016-NA-NA
  #           2016-01-NA
  #           2016-01-01
  #           2016-01
}

##### Function to calculate the birthorder based on the siblings still alive at the time of birth
older_sibs_alive_and_dependent = function(byear, dyear) {
	sibs = length(byear)
	older_sibs_alive_and_dependent = integer(length=sibs) + 1
	for(i in 1:sibs) {
		older_sibs = byear <= byear[i] # not using < because of twins
		older_sibs[i] = F # minus self
		my_sibs = sum(older_sibs,na.rm = T) # minus self
		if(my_sibs > 0) {
			sib_births = byear[ which(older_sibs) ]
			sib_deaths = dyear[ which(older_sibs) ]
			my_sibs = my_sibs -
				sum(
					# sib_births < (byear[i] - 5) | # others born more than 5y earlier than me  # 10 seconds of 17
						(sib_deaths <= byear[i]) # died before my birth
				,na.rm=T)
			older_sibs_alive_and_dependent[i] = my_sibs
		}
	}
	older_sibs_alive_and_dependent
}

```



## Load data
```{r Load data}
### Informations about individuals living in the household in 2014/2015
## All Individuals living in the household
bk_ar1 = read_dta("data/hh14_all_dta/bk_ar1.dta") # Book K, Section ar
# compute father pidlink
bk_ar1 = left_join(bk_ar1, bk_ar1 %>% select(hhid14_9, pid14, pidlink) %>% rename(ar10 = pid14, father_pidlink = pidlink), by = c("hhid14_9", "ar10"))
# compute mother pidlink
bk_ar1 = left_join(bk_ar1, bk_ar1 %>% select(hhid14_9, pid14, pidlink) %>% rename(ar11 = pid14, mother_pidlink = pidlink), by = c("hhid14_9", "ar11"))

### Informations from IFLS wave 5 to link data to earlier waves:
ptrack = read_dta("data/hh14_all_dta/ptrack.dta") # Traking informations

### Pregnancy Informations from mother
## Wave 5 - 2014
w5_pregnancy = read_dta("data/hh14_all_dta/b4_ch1.dta") # Book 4, Section ch
## Wave 4 - 2007
w4_pregnancy = read_dta("data/hh07_all_dta/b4_ch1.dta") # Book 4, Section ch
## Wave 3 - 2000
w3_pregnancy = read_dta("data/hh00_all_dta/b4_ch1.dta") # Book 4, Section ch
## Wave 2 - 1997
w2_pregnancy = read_dta("data/hh97dta/b4_ch1.dta") # Book 4, Section ch
## Wave 1 - 1993
w1_pregnancy = read_dta("data/hh93dta/buk4ch1.dta") # Book 4, Section ch

### Marriage information from mother
## Wave 5 - 2014
w5_marriage= read_dta("data/hh14_all_dta/b4_kw3.dta") # Book 4, Section kw3
## Wave 4 - 2007
w4_marriage = read_dta("data/hh07_all_dta/b4_kw2.dta") # Book 4, Section kw2
## Wave 3 - 2000
w3_marriage = read_dta("data/hh00_all_dta/b4_kw3.dta") # Book 4, Section kw3
## Wave 2 - 1997
w2_marriage = read_dta("data/hh97dta/b4_kw2.dta") # Book 4, Section kw2
## Wave 1 - 1993
w1_marriage = read_dta("data/hh93dta/buk4kw2.dta") # Book 4, Section kw2

### IQ Information
ek_ek2 = read_dta("data/hh14_all_dta/ek_ek2.dta") # Book ek2
# additional information (counting backwards, adaptive testing) for adults
b3b_cob = read_dta("data/hh14_all_dta/b3b_cob.dta") # Book 3b, Section cob
b3b_co1 = read_dta("data/hh14_all_dta/b3b_co1.dta") # Book 3b, Section co1
### Personality Information (only for adults)
b3b_psn = read_dta("data/hh14_all_dta/b3b_psn.dta") # Book 3b, Section psn

### Risk taking
b3a_si = read_dta("data/hh14_all_dta/b3a_si.dta") # Book 3b, Section si

```


## Pregnancy informations
```{r}
## Select data
w5_pregnancy = w5_pregnancy %>% select(pidlink, ch06, ch08, ch09day, ch09mth, ch09yr, ch25)
w4_pregnancy = w4_pregnancy %>% select(pidlink, ch06, ch08, ch09day, ch09mth, ch09yr, ch25)
w3_pregnancy = w3_pregnancy %>% select(pidlink, ch06, ch08, ch09day, ch09mth, ch09yr, ch25)
w2_pregnancy = w2_pregnancy %>% select(pidlink, ch06, ch08, ch09day, ch09mth, ch09yr, ch25)
w1_pregnancy = w1_pregnancy %>% select(pidlink, ch06, ch08, ch09day, ch09mth, ch09yr, ch25)
# In the first wave the year is named wrong
w1_pregnancy = w1_pregnancy %>%
  filter(ch09yr <=93)
w1_pregnancy$ch09yr = as.numeric(str_c("19", w1_pregnancy$ch09yr))

## Combine data
pregnancy = bind_rows(w1_pregnancy, w2_pregnancy, w3_pregnancy, w4_pregnancy, w5_pregnancy)
length(unique(pregnancy$pidlink))

## Rename Variables
pregnancy = pregnancy %>% rename(lifebirths = ch06, gender = ch08, birth_day = ch09day, birth_month = ch09mth, birth_year = ch09yr, mother_pidlink = pidlink, alive = ch25) # pregnancy$lifebirths values: 1 = still pregnant, 2 = livebirth, 3 = still birth, 4 = misscarriage

## Set values as NA that are missing
pregnancy$birth_day[ pregnancy$birth_day>31] = NA
pregnancy$birth_month[ pregnancy$birth_month>12] = NA
pregnancy$birth_year[ pregnancy$birth_year>2016] = NA
pregnancy$birth_day[ is.nan(pregnancy$birth_day) ] = NA
pregnancy$birth_month[ is.nan(pregnancy$birth_month) ] = NA
pregnancy$birth_year[ is.nan(pregnancy$birth_year)] = NA


pregnancy = pregnancy %>%
  mutate(birthdate = all_available_info_birth_date(birth_year, birth_month, birth_day),
         mother_birthdate = str_c(mother_pidlink, "-", birthdate))

##remove all with missing birthdate/miscarriage date
pregnancy = pregnancy %>%
  filter(!is.na(birthdate))

pregnancy_alive = pregnancy %>%
  filter(lifebirths ==2)

### Remove all families with multiple births:
## Mark multpile births
pregnancy = pregnancy %>%
  mutate(multiple_birth = ifelse(duplicated(mother_birthdate), 1, 0))
## All women that ever had multiple births
twinmothers = pregnancy %>%
  filter(multiple_birth == 1)
length(unique(twinmothers$mother_pidlink)) ## number of families that have to be excluded due to multiple births

pregnancy = pregnancy %>%
    filter(!(mother_pidlink %in% twinmothers$mother_pidlink))


### Remove all families with multiple births:
## Mark multpile births
pregnancy_alive = pregnancy_alive %>%
  mutate(multiple_birth = ifelse(duplicated(mother_birthdate), 1, 0))
## All women that ever had multiple births
twinmothers_alive = pregnancy_alive %>%
  filter(multiple_birth == 1)
length(unique(twinmothers_alive$mother_pidlink)) ## number of families that have to be excluded due to multiple births

pregnancy_alive = pregnancy_alive %>%
    filter(!(mother_pidlink %in% twinmothers$mother_pidlink))

```


## Uterus Birthorder and maternal sibling count
```{r Uterus birthorder and maternal sibling count}
### Birthorder calculations
## Biological Uterus Birthorder calculations with all pregnancies:
pregnancy1 = pregnancy %>%
  group_by(mother_pidlink) %>%
  mutate(birthorder_uterus_preg = min_rank(birthdate),
         sibling_count_uterus_preg = sum(!is.na(birthdate)))
         
## Biological Uterus Birthorder calculations with all births:
pregnancy2 = pregnancy %>%
  filter(lifebirths == 2) %>%
  group_by(mother_pidlink) %>%
    mutate(birthorder_uterus_alive = min_rank(birthdate),
           sibling_count_uterus_alive = sum(!is.na(birthdate)))

pregnancy2 = pregnancy2 %>% select(mother_birthdate, birthorder_uterus_alive, sibling_count_uterus_alive)

### 3040 individuals are removed because of death before/during birth

### Combine birthorder data
pregnancy = left_join(pregnancy1, pregnancy2, by="mother_birthdate")

### How many no-only-children are left?
pregnancy = as.data.frame(pregnancy)
pregnancy$mother_pidlink = pregnancy$mother_pidlink.y
pregnancy$mother_pidlink.y = pregnancy$mother_pidlink.x = NULL

pregnancy_check = pregnancy %>%
  filter(sibling_count_uterus_preg != 1)

pregnancy_alive_check = pregnancy %>%
  filter(!is.na(sibling_count_uterus_alive), sibling_count_uterus_alive != 1)


### How many participants are still alive by IFLS 5:
pregnancy_alive_check = pregnancy_alive_check %>%
  filter(alive != 3)

### How many participants are 15 or older by the time of ifls 5?
pregnancy_alive_and_old_enough_check = pregnancy_alive_check %>%
    filter(birth_year < 2000)



### Graphs
## Biological Birthorder - Uterus_Pregnancies
qplot(pregnancy$birthorder_uterus_preg)
ggplot(pregnancy, aes(x=sibling_count_uterus_preg, y=birthorder_uterus_preg)) + geom_jitter()

## Biological Birthorder - Uterus_Births
qplot(pregnancy$birthorder_uterus_alive)
ggplot(pregnancy, aes(x=sibling_count_uterus_alive, y=birthorder_uterus_alive)) + geom_jitter()

## Bio: Uterus_preg vs. Uterus_Births
ggplot(pregnancy, aes(x=birthorder_uterus_preg, y=birthorder_uterus_alive)) + geom_jitter()
# The birth_order_alive is always lower, which makes sense, becaus not live births (miscarriage, still births are excluded)


```

## Marriage History
```{r marriage history}
## Select data
w5_marriage = w5_marriage %>% select(pidlink, kw10mth, kw10yr, kw18mth, kw18yr)
w4_marriage = w4_marriage %>% select(pidlink, kw10mth, kw10yr, kw18mth, kw18yr)
w3_marriage = w3_marriage %>% select(pidlink, kw10mth, kw10yr, kw18mth, kw18yr)
w2_marriage = w2_marriage %>% select(pidlink, kw10mth, kw10yr, kw18mth, kw18yr)
w1_marriage = w1_marriage %>% select(pidlink, kw05a, kw05b, kw13a, kw13b)
# In the first wave the year is named wrong
w1_marriage = w1_marriage %>%
  mutate(kw05a = ifelse(kw05a <= 93, as.numeric(str_c("19", w1_marriage$kw05a)), kw05a),
         kw13a = ifelse(kw13a <=93 , as.numeric(str_c("19", w1_marriage$kw13a)), kw13a))
# And the column names are wrong...
w1_marriage = w1_marriage %>% rename(kw10mth = kw05b, kw10yr = kw05a, kw18mth = kw13b, kw18yr = kw13a)

## Combine marriage informations
marriage = bind_rows(w1_marriage, w2_marriage, w3_marriage, w4_marriage, w5_marriage)
# Rename columns
marriage = marriage %>% rename(start_year = kw10yr, start_month = kw10mth, end_year = kw18yr, end_month = kw18mth)
marriage = marriage %>% arrange(pidlink, start_year, start_month, end_year, end_month)
marriage = marriage %>% filter(!duplicated(cbind(pidlink, start_year, start_month)) | !is.na(start_year) | is.na(start_month)) # nobody gets married twice on the same day, right? so these are dupes.
# Set values as NA that are missing
marriage$start_year[ marriage$start_year<1900] = NA
marriage$start_year[ marriage$start_year>2016] = NA
marriage$start_year[ is.nan(marriage$start_year)] = NA
marriage$end_year[ marriage$end_year<1900] = NA
marriage$end_year[ marriage$end_year>2016] = NA
marriage$end_year[ is.nan(marriage$end_year)] = NA
marriage$start_month [marriage$start_month>12] = NA
marriage$start_month [is.nan(marriage$start_month)] = NA
marriage$end_month [marriage$end_month>12] = NA
marriage$end_month [is.nan(marriage$end_month)] = NA

table(is.na(marriage$start_year))
table(is.na(marriage$end_year))
library(lubridate)
## Calculate date for beginning of marriage:
pad_month = function(x) { str_pad(x, width = 2, side = "left", pad = "0")}
marriage = marriage %>% 
  mutate(
        start_string = str_c(start_year, "-", ifelse(is.na(start_month), "01", pad_month(start_month)), "-01"),
        end_string = str_c(end_year, "-", ifelse(is.na(end_month), "12", pad_month(end_month)), "-01")
  )
marriage = marriage %>% ungroup() %>% mutate(start = ymd(start_string),
                               end = ymd(end_string) + months(1) - days(1))

# marriage$start_string[!is.na(marriage$start_string) & is.na(marriage$start) ]
# marriage$end_string[!is.na(marriage$end_string) & is.na(marriage$end) ]
# 
qplot(marriage$start)
qplot(marriage$end)
## Count number of marriages
marriage = marriage %>%
  arrange(pidlink, start, end) %>%
  group_by(pidlink) %>%
  mutate(number_marriages = n(),
         order_marriage = row_number()) %>%
  mutate(marriage_id = paste0(pidlink, "_", as.character(start), "/",as.character(end)))

dupes = unique(marriage$marriage_id[duplicated(marriage$marriage_id)])

## Remove all families with unclear fathers
marriage = marriage %>%
  filter(! marriage_id %in% dupes)

## Graphic analysis
marriage %>% ggplot(aes(number_marriages, order_marriage)) + geom_jitter(width=0.1)

## Problem: What do i do with still ongoing marriages? How do i deal with marriages with missing end, but ther are other marriages afterwards (can i use that start date as end date? and if yes --> how?!?)

table(is.na(marriage$start))
table(is.na(marriage$end))

pregnancy$month = paste0(pregnancy$birth_year,"-", ifelse(is.na(pregnancy$birth_month), "01", pad_month(pregnancy$birth_month)))

minimum_start = min(ymd(str_c(pregnancy$month, "-01")), na.rm = T)
maximum_end = max(ymd(str_c(pregnancy$month, "-01")), na.rm = T)
marriage_timeline = marriage %>%
    mutate(implied_start = as.Date(ifelse(is.na(start), minimum_start , start), origin="1970-01-01"),
           implied_end = as.Date(ifelse(is.na(end), maximum_end , end), origin="1970-01-01")) %>%
  filter(start < end)

marriage_timeline = marriage_timeline %>%
    rowwise() %>%
    do(data.frame(
        marriage_id=.$marriage_id, 
        mother_pidlink = .$pidlink,
        order_marriage = .$order_marriage,
        start = .$start,
        end = .$end,
        month=seq(.$implied_start,.$implied_end,by="1 month")))

marriage_timeline = marriage_timeline %>% 
  arrange(mother_pidlink, start, end) %>% 
  distinct(mother_pidlink, month, .keep_all = TRUE)


marriage_timeline$month = stringr::str_sub(as.character(marriage_timeline$month),1,7)

length(intersect(pregnancy$mother_pidlink, marriage_timeline$mother_pidlink))
length(setdiff(pregnancy$mother_pidlink, marriage_timeline$mother_pidlink))
length(setdiff(marriage_timeline$mother_pidlink, pregnancy$mother_pidlink))

pregnancy = pregnancy %>% left_join(marriage_timeline, by = c("mother_pidlink", "month"))

rm(marriage_timeline)

table(table(pregnancy$marriage_id))
sum(table(table(pregnancy$marriage_id)))
```




## Select individual data
```{r select individual data}
### Individuals
individuals = bk_ar1 %>% select(hhid14_9, pidlink, father_pidlink, mother_pidlink, ar01a, ar02b, ar10, ar11, ar07, ar08day, ar08mth, ar08yr, ar09, ar18eyr, ar18emth)
#Rename variables to make it easier
individuals = rename(individuals, relation_to_HH_head = ar02b, fatherID = ar10, motherID = ar11, sex = ar07, age = ar09, status = ar01a, death_yr = ar18eyr, death_month = ar18emth) 
# Remove duplicats (some people are mentioned in two households, e.g. because they moved in the last 12 months)
individuals = individuals %>% distinct(pidlink, .keep_all = TRUE)
individuals_unchanged = individuals

## people whose parents can not be identified have to be marked as NA:
individuals$fatherID[ individuals$fatherID>50] = NA
individuals$motherID[ individuals$motherID>50] = NA


## create mother_father_number that should be identical for all siblings in one family that share the same parents
individuals = individuals %>% mutate(
  mother_father_dyad = stringr::str_c(father_pidlink, mother_pidlink)
) # father_pidlink and mother_pidlink; if one is missing = NA


## Create date of birth
#Set all variables missing that have not been reported:
individuals$ar08day[ individuals$ar08day>31] = NA
individuals$ar08mth[ individuals$ar08mth>12] = NA
individuals$ar08yr[ individuals$ar08yr>2016] = NA
individuals$ar08day[ is.nan(individuals$ar08day) ] = NA
individuals$ar08mth[ is.nan(individuals$ar08mth) ] = NA
individuals$ar08yr[ is.nan(individuals$ar08yr)] = NA
individuals$death_month[ individuals$death_month>12] = NA
individuals$death_yr[ individuals$death_yr>2016] = NA
individuals$death_month[ is.nan(individuals$death_month) ] = NA
individuals$death_yr[ is.nan(individuals$death_yr)] = NA

## Create variable that contains pidlink of mother and birthdate of child:
individuals = individuals %>%
    mutate(mother_birthdate = str_c(mother_pidlink, "-",
          all_available_info_birth_date(ar08yr, ar08mth, ar08day))) # mother_pidlink-YYYY-MM; is NA if birth_year is missing

##Remove all with missing mother_birthdate
individuals = individuals %>%
  filter(!is.na(mother_birthdate))


### Remove all families with multiple births:
## Mark multpile births
individuals = individuals %>%
  mutate(multiple_birth = ifelse(duplicated(mother_birthdate), 1, 0))

## All women that ever had multiple births
twinmothers_individual = individuals %>%
  filter(multiple_birth == 1)
length(unique(twinmothers_individual$mother_pidlink)) # number of families that have to be excluded due to multiple birth

individuals = individuals %>%
    filter(!(mother_pidlink %in% twinmothers_individual$mother_pidlink))

```



## Combine informations from pregnancy file and individual file
```{r Combine informations from pregnancy file and individual file}
## Combine informations from pregnancy history and individual files
alldata_pregnancy = inner_join(individuals, pregnancy, by = "mother_birthdate")
alldata_pregnancy_check = inner_join(individuals, pregnancy_alive_and_old_enough_check, by = "mother_birthdate")
```

## Gene birthorder (based on both parents)
```{r gene birthorder}
### Gene birthorder is calculated by grouping individuals together, for whom we know both parents and making a rank list
alldata_pregnancy = alldata_pregnancy %>%
  group_by(mother_father_dyad) %>%
    mutate(birthorder_genes = min_rank(birthdate),
           birthorder_genes = ifelse(is.na(mother_father_dyad), NA, birthorder_genes),
           sibling_count_genes = sum(!is.na(mother_father_dyad)))
ggplot(alldata_pregnancy, aes(x=sibling_count_genes, y=birthorder_genes)) + geom_jitter()

## Plot uterus birth order vs. gene birthorder
ggplot(alldata_pregnancy, aes(x=birthorder_uterus_alive, y=birthorder_genes)) + geom_jitter()

```

## Select IQ data
```{r select IQ data}
### IQ Informations
##ek2 (>14yrs)
iq2.1 = ek_ek2 %>% select(hhid14_9, pidlink, age, sex, ektype, resptype, result, reason, ek1_ans, ek2_ans, ek3_ans, ek4_ans, ek5_ans, ek6_ans, ek7_ans, ek8_ans, ek9_ans, ek10_ans, ek11_ans, ek12_ans, ek13_ans, ek14_ans, ek15_ans, ek16_ans, ek17_ans, ek18_ans, ek19_ans, ek20_ans, ek21_ans, ek22_ans)


##additional informations for adults: counting backwards
iq2.2 = b3b_co1 %>% select(hhid14_9, pidlink, co04a, co04b, co04c, co04d, co04e, co07count, co10count)
##additional informations for adults: adaptive number test
iq2.3 = b3b_cob %>% select(hhid14_9, pidlink, w_abil, cob18, cob19b)

## put all the informations for participants >= 15 together
iq2 = full_join(iq2.1, iq2.2, by = "pidlink")
iq2 = full_join(iq2, iq2.3, by = "pidlink")
iq = iq2
iq <- plyr::rename(iq, c("age"="IQage")) 

### calculate iq scores
##Raven Test
answered_raven_items = iq %>% select(ek1_ans:ek6_ans, ek11_ans, ek12_ans)
psych::alpha(data.frame(answered_raven_items))
iq$raven = rowMeans( answered_raven_items, na.rm = T)
iq$raven[! iq$result %in% 1:2] = NA
qplot(iq$raven)

##Math Test
answered_math_items = iq %>% select(ek18_ans:ek22_ans)
psych::alpha(data.frame(answered_math_items))
iq$math = rowMeans( answered_math_items, na.rm = T)
iq$math[! iq$result %in% 1:2] = NA
qplot(iq$math)

##Counting Items
# Create Right/Wrong Scores for the counting items
iq$co04aright = as.numeric(iq$co04a == 93)
iq$co04bright = as.numeric(iq$co04b == iq$co04a-7)
iq$co04cright = as.numeric(iq$co04c == iq$co04b-7)
iq$co04dright = as.numeric(iq$co04d == iq$co04c-7)
iq$co04eright = as.numeric(iq$co04e == iq$co04d-7)

answered_counting_items = iq %>% select(co04aright:co04eright)
psych::alpha(data.frame(answered_counting_items))
iq$count_backwards = rowSums( answered_counting_items, na.rm = T) / 5
qplot(iq$count_backwards)

## Word Memory
iq$words_immediate = iq$co07count
iq$words_delayed = iq$co10count
qplot(iq$words_immediate, iq$words_delayed, geom = "jitter")
answered_word_items = iq %>% select(co07count,co10count)
psych::alpha(data.frame(answered_word_items))
iq$words_remembered_avg = rowMeans( answered_word_items, na.rm = T)
qplot(iq$words_remembered_avg)

##Adaptive Numbering
iq$adaptive_numbering = iq$w_abil


#iq$drew_pentagons = as.numeric(iq$cob19b == 1)
#iq$number_of_animals = iq$cob18

## Correlation of all Iq-Tests
round(cor(iq %>% select(raven, math, count_backwards, words_immediate, words_delayed, adaptive_numbering), use = "pairwise.complete.obs"), 2)

##Missingness_Patterns
formr::missingness_patterns(iq %>% select(raven, math, count_backwards, words_immediate, words_delayed, adaptive_numbering))

```

## compute g factor
```{r}

iq$words_delayed = iq$words_delayed/10
iq$words_immediate = iq$words_immediate/10
iq$adaptive_numbering = iq$adaptive_numbering/100


library(psych)
fa.parallel(iq %>% select(raven, math, count_backwards, words_delayed, adaptive_numbering) %>% data.frame())
fa(iq %>% select(raven, math, count_backwards, words_delayed, adaptive_numbering) %>% data.frame())
omega(iq %>% select(raven, math, count_backwards, words_delayed, adaptive_numbering) %>% data.frame())

library(lavaan)
"g_factor =~ raven + math + count_backwards +  words_delayed+ adaptive_numbering" %>%
  cfa(missing = "fiml", data = iq) -> cfa_g
summary(cfa_g)

iq$g_factor = predict(cfa_g)[,1]
qplot(iq$g_factor)
round(cor(iq %>% select(g_factor, raven, math, count_backwards, words_immediate, words_delayed, adaptive_numbering), use = "pairwise.complete.obs"), 2)

formr::missingness_patterns(iq %>% select(raven, math, count_backwards, words_immediate, words_delayed, adaptive_numbering, g_factor)) ## oops, lavaan only predicts the g factor for complete cases

# impute missing values
# library(missMDA)
# estimate number of components
# rr = iq %>% select(raven, math, count_backwards, words_immediate, words_delayed, adaptive_numbering)
# nb <- estim_ncpPCA(rr, ncp.min=0, ncp.max=5)
# actual impute
# rr.impute <- imputePCA(rr, ncp=nb$ncp)
# https://stats.stackexchange.com/questions/123725/how-can-i-estimate-a-principal-component-from-incomplete-data?noredirect=1&lq=1

# Run pca
# pca.fit <- prcomp(rr.impute$completeObs)
# pca.fit$rotation
```


## Select personality data
```{r select pesonality data}
### Personality
##Rearrange personality data so that every individual has only one row
pers = spread(b3b_psn, psntype, psn01)
##name columns
colnames(pers) <- c("hhid14_9", "pid14", "hhid14", "pidlink", "version", "module", "e1", "c1", "o1", "e2r", "n1r", "a1", "n2r", "o2", "c2r", "o3", "a2", "c3", "e3", "a3r", "n3")
pers = pers %>% select(hhid14_9, pidlink, e1, c1, o1, e2r, n1r, a1, n2r, o2, c2r, o3, a2, c3, e3, a3r, n3)

##Extraversion
pers$e2r_reversed = 6 - pers$e2r
extraversion = pers %>% select(e1, e2r_reversed, e3)
psych::alpha(data.frame(extraversion), check.keys = T)
pers$big5_ext = rowMeans(extraversion)
qplot(pers$big5_ext)

##conscientiousness
pers$c2r_reversed = 6 - pers$c2r
conscientiousness = pers %>% select(c1, c2r_reversed, c3)
psych::alpha(data.frame(conscientiousness), check.keys = T)
pers$big5_con = rowMeans(conscientiousness)
qplot(pers$big5_con)

##Openness
openness = pers %>% select(o1, o2, o3)
psych::alpha(data.frame(openness), check.keys = T)
pers$big5_open = rowMeans(openness)
qplot(pers$big5_open)

## Neuroticism
pers$n1r_reversed = 6 - pers$n1r
pers$n2r_reversed = 6 - pers$n2r
neuroticism = pers %>% select(n1r_reversed, n2r_reversed, n3)
pers$big5_neu = rowMeans(neuroticism)
qplot(pers$big5_neu)

##Agreeableness
pers$a3r_reversed = 6- pers$a3r
agreeableness= pers %>% select(a1, a2, a3r_reversed)
pers$big5_agree = rowMeans(agreeableness)
qplot(pers$big5_agree)
```


## Select risk taking data
```{r select risk taking data}
###Risktaking
risk = b3a_si %>% select(hhid14_9, pidlink, random_si, si01, si02, si03, si04, si05, si11, si12, si13, si14, si15)

## 8 means they didnt know which answer they would choose
risk$si01[ risk$si01 == 8] = NA
risk$si02[ risk$si02 == 8] = NA
risk$si03[ risk$si03 == 8] = NA
risk$si04[ risk$si04 == 8] = NA
risk$si05[ risk$si05 == 8] = NA
risk$si11[ risk$si11 == 8] = NA
risk$si12[ risk$si12 == 8] = NA
risk$si13[ risk$si13 == 8] = NA
risk$si14[ risk$si14 == 8] = NA
risk$si15[ risk$si15 == 8] = NA

## calculate a risk score for risk game A 
# (5 = gamble averse, Ordinalskala : 1 = risk loving, 4 = risk averse)
risk$riskA = ifelse(risk$si01 == 1 & risk$si02 == 1, 5,
             ifelse(risk$si01 == 2 & risk$si03 == 1 & risk$si04 == 1, 4,
             ifelse(risk$si01 == 2 & risk$si03 == 1 & risk$si04 == 2, 3,
             ifelse(risk$si01 == 2 & risk$si03 == 2 & risk$si05 == 1, 2,
             ifelse(risk$si01 == 2 & risk$si03 == 2 & risk$si05 == 2, 1,
             NA)))))
qplot(risk$riskA[risk$riskA != 5])

## calculate a risk score for risk game B 
# (5 = gamble averse, Ordinalskala : 1 = risk loving, 4 = risk averse)
risk$riskB = ifelse(risk$si11 == 2 & risk$si12 == 1, 5,
             ifelse(risk$si11 == 1 & risk$si13 == 1 & risk$si14 == 1, 4,
             ifelse(risk$si11 == 1 & risk$si13 == 1 & risk$si14 == 2, 3,
             ifelse(risk$si11 == 1 & risk$si13 == 2 & risk$si15 == 1, 2,
             ifelse(risk$si11 == 1 & risk$si13 == 2 & risk$si15 == 2, 1,
             NA)))))
qplot(risk$riskB[risk$riskB !=5])

psych::alpha(data.frame(risk %>% select(riskA, riskB)), check.keys = T)


```


## Merge data

```{r merge data}
### Merge all data for people with birthorder informations
alldata_birthorder = left_join(alldata_pregnancy, iq, by = "pidlink")
alldata_birthorder = left_join(alldata_birthorder, pers, by = "pidlink")
alldata_birthorder = left_join(alldata_birthorder, risk, by = "pidlink")

alldata = left_join(individuals_unchanged, iq, by = "pidlink")
alldata = left_join(alldata, pers, by = "pidlink")
alldata = left_join(alldata, risk, by = "pidlink")

alldata_check = left_join(alldata_pregnancy_check, iq, by = "pidlink")
alldata_check = left_join(alldata_check, pers, by = "pidlink")
alldata_check = left_join(alldata_check, risk, by = "pidlink")


### Which values are missing?
## All data
formr::missingness_patterns(alldata_check %>% select(raven, math, count_backwards, words_immediate, words_delayed, adaptive_numbering, big5_ext, riskA, riskB))
## Intelligence
formr::missingness_patterns(alldata_check %>% select(raven, math, count_backwards, words_immediate, words_delayed, adaptive_numbering))
##Personality
formr::missingness_patterns(alldata_check %>% select(big5_ext, big5_con, big5_open, big5_neu, big5_agree))
##Risk
formr::missingness_patterns(alldata_check %>% select(riskA, riskB))
```


## Exploring data

``` {r exploring data}
### Personality
ggplot(alldata, aes(x=big5_agree, y=big5_neu)) + geom_jitter()

###Openness and Intelligence
round(cor(alldata %>% ungroup %>% select(raven, math, count_backwards, words_immediate, words_delayed, adaptive_numbering, g_factor, big5_open), use = "pairwise.complete.obs"), 2)

###Other personality factors and Intelligence
round(cor(alldata %>% ungroup %>% select(raven, math, count_backwards, words_immediate, words_delayed, adaptive_numbering, g_factor, big5_con), use = "pairwise.complete.obs"), 2)
round(cor(alldata %>% ungroup %>% select(raven, math, count_backwards, words_immediate, words_delayed, adaptive_numbering, g_factor, big5_ext), use = "pairwise.complete.obs"), 2)
round(cor(alldata %>% ungroup %>% select(raven, math, count_backwards, words_immediate, words_delayed, adaptive_numbering, g_factor, big5_agree), use = "pairwise.complete.obs"), 2)
round(cor(alldata %>% ungroup %>% select(raven, math, count_backwards, words_immediate, words_delayed, adaptive_numbering, g_factor, big5_neu), use = "pairwise.complete.obs"), 2)
# Opennes has the highest values!
ggplot(alldata, aes(x=big5_open, y=g_factor)) + geom_jitter() + geom_smooth(method=lm) 

### Birth Order
alldata_birthorder %>%
  filter(birthorder_uterus_alive<6) %>%
  ggplot(aes(x=birthorder_uterus_alive, y=g_factor)) + geom_jitter() + geom_smooth(method=lm) 

table(alldata_birthorder$sibling_count_uterus_alive)

alldata_birthorder %>%
  filter(birthorder_uterus_alive<8) %>%
  filter(sibling_count_uterus_alive<8) %>%
  ggplot(aes(x=birthorder_uterus_alive, y=g_factor)) + 
  facet_wrap(~ sibling_count_uterus_alive) + 
  geom_jitter() + geom_smooth(method=lm) 

alldata_birthorder %>% select(birthorder_uterus_alive, sibling_count_uterus_alive, birthdate) -> subs
alldata_birthorder$byear = as.numeric(stringr::str_sub(alldata_birthorder$birthdate,1,4))

alldata_birthorder %>%
  filter(birthorder_uterus_alive < 8) %>%
  filter(sibling_count_uterus_alive < 8) %>%
   mutate(birthorder_uterus_alive = factor(birthorder_uterus_alive, levels = c("1","2","3","4","5","6", "7", "8")),
         sibling_count_uters_alive = factor(sibling_count_uterus_alive)) %>%
  ggplot(aes(x=birthorder_uterus_alive, y=g_factor, colour = sibling_count_uterus_alive, group = sibling_count_uterus_alive)) + 
  geom_pointrange(fun.data = "mean_cl_boot", stat = "summary", position = position_dodge(width = 0.4)) + 
  geom_line(fun.data = "mean_cl_boot", stat = "summary", position = position_dodge(width = 0.4))

alldata_birthorder %>%
  filter(birthorder_uterus_alive < 8) %>%
  filter(sibling_count_uterus_alive < 8) %>%
   mutate(birthorder_uterus_alive = factor(birthorder_uterus_alive,levels = c("1","2","3","4","5","6", "7", "8")),
         sibling_count_uters_alive = sibling_count_uterus_alive) %>%
  ggplot(aes(x=birthorder_uterus_alive, y=big5_open, colour = sibling_count_uterus_alive, group = sibling_count_uterus_alive)) + 
  geom_pointrange(fun.data = "mean_cl_boot", stat = "summary", position = position_dodge(width = 0.4)) + 
  geom_line(fun.data = "mean_cl_boot", stat = "summary", position = position_dodge(width = 0.4))

alldata_birthorder %>%
  filter(birthorder_uterus_alive < 8) %>%
  filter(sibling_count_uterus_alive < 8) %>%
   mutate(birthorder_uterus_alive = factor(birthorder_uterus_alive, levels = c("1","2","3","4","5","6", "7", "8")),
         sibling_count_uters_alive = sibling_count_uterus_alive) %>%
  ggplot(aes(x=birthorder_uterus_alive, y=big5_ext, colour = sibling_count_uterus_alive, group = sibling_count_uterus_alive)) + 
  geom_pointrange(fun.data = "mean_cl_boot", stat = "summary", position = position_dodge(width = 0.4)) + 
  geom_line(fun.data = "mean_cl_boot", stat = "summary", position = position_dodge(width = 0.4))


### Genes:
alldata_birthorder %>%
  filter(birthorder_genes < 8) %>%
  filter(sibling_count_genes < 8) %>%
   mutate(birthorder_genes = factor(birthorder_genes),
         sibling_count_genes = sibling_count_genes) %>%
  ggplot(aes(x=birthorder_genes, y=g_factor, colour = sibling_count_genes, group = sibling_count_genes)) + 
  geom_pointrange(fun.data = "mean_cl_boot", stat = "summary", position = position_dodge(width = 0.4)) + 
  geom_line(fun.data = "mean_cl_boot", stat = "summary", position = position_dodge(width = 0.4))


alldata_birthorder %>%
  filter(birthorder_genes < 8) %>%
  filter(sibling_count_genes < 8) %>%
   mutate(birthorder_genes = factor(birthorder_genes),
         sibling_count_genes = sibling_count_genes) %>%
  ggplot(aes(x=birthorder_genes, y=big5_open, colour = sibling_count_genes, group = sibling_count_genes)) + 
  geom_pointrange(fun.data = "mean_cl_boot", stat = "summary", position = position_dodge(width = 0.4)) + 
  geom_line(fun.data = "mean_cl_boot", stat = "summary", position = position_dodge(width = 0.4))

alldata_birthorder %>%
  filter(birthorder_uterus_alive < 8) %>%
  filter(sibling_count_uterus_alive < 8) %>%
   mutate(birthorder_uterus_alive = factor(birthorder_uterus_alive, levels = 1:7),
         sibling_count_uters_alive = factor(sibling_count_uterus_alive, levels = 1:7)) %>%
  ggplot(aes(x=birthorder_uterus_alive, y=big5_ext, colour = sibling_count_uterus_alive, group = sibling_count_uterus_alive)) + 
  geom_pointrange(fun.data = "mean_cl_boot", stat = "summary", position = position_dodge(width = 0.4)) + 
  geom_line(fun.data = "mean_cl_boot", stat = "summary", position = position_dodge(width = 0.4))

```


## Save data

for future analyses

```{r save data}
saveRDS(alldata_birthorder, file = "data/alldata_birthorder.rds")

saveRDS(alldata, file = "data/alldata.rds")

```
