# Inclusion Process {.tabset}

## Helper
```{r helper}
source("0_helpers.R")
```

## Load data
```{r Load data}
### Import alldata used for birthorder
alldata_birthorder = readRDS("data/alldata_birthorder.rds")

### Import all alldata
alldata = readRDS("data/alldata.rds")
```

## Birthorder
### Total
Number of individuals we were able to reconstruct the three different birth orders based on pregnancy and marriage history for 
```{r}
## Count Birthorder
n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate)),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate)),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate)))
print(n_birthorder)
```


### No Multiple Births
``` {r}
n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1))
print(n_birthorder)

```

### No only Children
```{r}
## Count Birthorder - without only children
n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1))
print(n_birthorder)
```


### Still alive
```{r}
## Count Birthorder - without only children
n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1 &
                                         alive != 3, na.rm = T),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1 & alive != 3, na.rm = T),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1 & alive != 3, na.rm =T))
print(n_birthorder)
```

### Aged 15 or older
```{r}
## Count Birthorder - without only children
n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1 &
                                         alive != 3 & birth_year <= 2000, na.rm = T),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1 & alive != 3 & birth_year <= 2000,
                                           na.rm = T),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1 & alive != 3 & birth_year <= 2000,
                                                    na.rm =T))
print(n_birthorder)
```


### Interviewed for IFLS 5  {.tabset}

#### Total
```{r}
### All people in the birthorder data file that have ANY g-factor, personality or risk values have to be marked
alldata_birthorder = alldata_birthorder %>%
  mutate(used_all = ifelse(!is.na(g_factor) || !is.na(big5_ext) || !is.na(big5_open) || !is.na(big5_agree) ||
                             !is.na(big5_neu) || !is.na(big5_con) || !is.na(riskA) || !is.na(riskB), 1, 0))

n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1 &
                                         alive != 3 & birth_year <= 2000 & used_all == 1, na.rm = T),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1 & alive != 3 & birth_year <= 2000 &
                                             used_all == 1,
                                           na.rm = T),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1 & alive != 3 & birth_year <= 2000 &
                                                      used_all == 1,
                                                    na.rm =T))
print(n_birthorder)
```

#### Intelligence
```{r}
### All people in the birthorder data file that have all Intelligence measurements
alldata_birthorder = alldata_birthorder %>%
  mutate(used_iq = ifelse(!is.na(g_factor), 1, 0)) # used_iq = 1:included, 0:not_included

n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1 &
                                         alive != 3 & birth_year <= 2000 & used_iq == 1, na.rm = T),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1 & alive != 3 & birth_year <= 2000 &
                                             used_iq == 1,
                                           na.rm = T),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1 & alive != 3 & birth_year <= 2000 &
                                                      used_iq == 1,
                                                    na.rm =T))
print(n_birthorder)

```

#### Personality
```{r}
### All people in the birthorder data file that have Personality Measurements
missingness_patterns(alldata_birthorder %>% select(big5_ext, big5_con, big5_open, big5_neu, big5_agree, big5_open))
# Missingness Pattern shows that if one of the personality measurements is missing all Measurements are missing
alldata_birthorder = alldata_birthorder %>%
  mutate(used_pers = ifelse(!is.na(big5_ext), 1, 0))
n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1 &
                                         alive != 3 & birth_year <= 2000 & used_pers == 1, na.rm = T),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1 & alive != 3 & birth_year <= 2000 &
                                             used_pers == 1,
                                           na.rm = T),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1 & alive != 3 & birth_year <= 2000 &
                                                      used_pers == 1,
                                                    na.rm =T))
print(n_birthorder)
```

#### Risk
```{r}
### All people in the birthorder data file that have Risk Measurements
missingness_patterns(alldata_birthorder %>% select(riskA, riskB))
## Missingness Pattern shows that sometimes A and sometimes B is missing
alldata_birthorder = alldata_birthorder %>%
  mutate(used_riskA = ifelse(!is.na(riskA), 1, 0),
         used_riskB = ifelse(!is.na(riskB), 1, 0))
n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1 &
                                         alive != 3 & birth_year <= 2000 & used_riskA == 1, na.rm = T),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1 & alive != 3 & birth_year <= 2000 &
                                             used_riskA == 1,
                                           na.rm = T),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1 & alive != 3 & birth_year <= 2000 &
                                                      used_riskA == 1,
                                                    na.rm =T))
print(n_birthorder)

n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1 &
                                         alive != 3 & birth_year <= 2000 & used_riskB == 1, na.rm = T),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1 & alive != 3 & birth_year <= 2000 &
                                             used_riskB == 1,
                                           na.rm = T),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1 & alive != 3 & birth_year <= 2000 &
                                                      used_riskB == 1,
                                                    na.rm =T))
print(n_birthorder)

```

# Inclusion Process {.tabset}

## Helper
```{r helper}
source("0_helpers.R")
```

## Load data
```{r Load data}
### Import alldata used for birthorder
alldata_birthorder = readRDS("data/alldata_birthorder.rds")

### Import all alldata
alldata = readRDS("data/alldata.rds")
```

## Birthorder
### Total
Number of individuals we were able to reconstruct the three different birth orders based on pregnancy and marriage history for 
```{r}
## Count Birthorder
n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate)),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate)),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate)))
print(n_birthorder)
```


### No Multiple Births
``` {r}
n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1))
print(n_birthorder)

```

### No only Children
```{r}
## Count Birthorder - without only children
n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1))
print(n_birthorder)
```


### Still alive
```{r}
## Count Birthorder - without only children
n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1 &
                                         alive != 3, na.rm = T),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1 & alive != 3, na.rm = T),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1 & alive != 3, na.rm =T))
print(n_birthorder)
```

### Aged 15 or older
```{r}
## Count Birthorder - without only children
n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1 &
                                         alive != 3 & birth_year <= 2000, na.rm = T),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1 & alive != 3 & birth_year <= 2000,
                                           na.rm = T),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1 & alive != 3 & birth_year <= 2000,
                                                    na.rm =T))
print(n_birthorder)
```


### Interviewed for IFLS 5  {.tabset}

#### Total
```{r}
### All people in the birthorder data file that have ANY g-factor, personality or risk values have to be marked
alldata_birthorder = alldata_birthorder %>%
  mutate(used_all = ifelse(!is.na(g_factor) || !is.na(big5_ext) || !is.na(big5_open) || !is.na(big5_agree) ||
                             !is.na(big5_neu) || !is.na(big5_con) || !is.na(riskA) || !is.na(riskB), 1, 0))

n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1 &
                                         alive != 3 & birth_year <= 2000 & used_all == 1, na.rm = T),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1 & alive != 3 & birth_year <= 2000 &
                                             used_all == 1,
                                           na.rm = T),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1 & alive != 3 & birth_year <= 2000 &
                                                      used_all == 1,
                                                    na.rm =T))
print(n_birthorder)
```

#### Intelligence
```{r}
### All people in the birthorder data file that have all Intelligence measurements
alldata_birthorder = alldata_birthorder %>%
  mutate(used_iq = ifelse(!is.na(g_factor), 1, 0)) # used_iq = 1:included, 0:not_included

n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1 &
                                         alive != 3 & birth_year <= 2000 & used_iq == 1, na.rm = T),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1 & alive != 3 & birth_year <= 2000 &
                                             used_iq == 1,
                                           na.rm = T),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1 & alive != 3 & birth_year <= 2000 &
                                                      used_iq == 1,
                                                    na.rm =T))
print(n_birthorder)

```

#### Personality
```{r}
### All people in the birthorder data file that have Personality Measurements
missingness_patterns(alldata_birthorder %>% select(big5_ext, big5_con, big5_open, big5_neu, big5_agree, big5_open))
# Missingness Pattern shows that if one of the personality measurements is missing all Measurements are missing
alldata_birthorder = alldata_birthorder %>%
  mutate(used_pers = ifelse(!is.na(big5_ext), 1, 0))
n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1 &
                                         alive != 3 & birth_year <= 2000 & used_pers == 1, na.rm = T),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1 & alive != 3 & birth_year <= 2000 &
                                             used_pers == 1,
                                           na.rm = T),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1 & alive != 3 & birth_year <= 2000 &
                                                      used_pers == 1,
                                                    na.rm =T))
print(n_birthorder)
```

#### Risk
```{r}
### All people in the birthorder data file that have Risk Measurements
missingness_patterns(alldata_birthorder %>% select(riskA, riskB))
## Missingness Pattern shows that sometimes A and sometimes B is missing
alldata_birthorder = alldata_birthorder %>%
  mutate(used_riskA = ifelse(!is.na(riskA), 1, 0),
         used_riskB = ifelse(!is.na(riskB), 1, 0))
n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1 &
                                         alive != 3 & birth_year <= 2000 & used_riskA == 1, na.rm = T),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1 & alive != 3 & birth_year <= 2000 &
                                             used_riskA == 1,
                                           na.rm = T),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1 & alive != 3 & birth_year <= 2000 &
                                                      used_riskA == 1,
                                                    na.rm =T))
print(n_birthorder)

n_birthorder = alldata_birthorder %>%
  summarise(Maternal_Birth_Order = sum(lifebirths == 2 & !is.na(mother_birthdate) &
                                         multiple_birth_in_family_alive != 1 & sibling_count_uterus_alive != 1 &
                                         alive != 3 & birth_year <= 2000 & used_riskB == 1, na.rm = T),
            Maternal_Pregnancy_Order = sum(!is.na(mother_birthdate) & multiple_birth_in_family_preg != 1 &
                                             sibling_count_uterus_preg != 1 & alive != 3 & birth_year <= 2000 &
                                             used_riskB == 1,
                                           na.rm = T),
            Parental_Full_Sibling_Birth_Order = sum(!is.na(marriage_id_birthdate) & 
                                                      multiple_birth_in_family_full != 1 & 
                                                      sibling_count_genes != 1 & alive != 3 & birth_year <= 2000 &
                                                      used_riskB == 1,
                                                    na.rm =T))
print(n_birthorder)

```



