---
output: html_document
editor_options: 
  chunk_output_type: console
---
# <span style="color:#A6D854">Birth Order Effects</span> {.tabset}

## Helper
```{r helper}
source("0_helpers.R")
```

## Load data
```{r Load Data}
birthorder = readRDS("data/alldata_birthorder.rds")
```

## Data preparations
```{r data preparations}
## For analyses we want to clean the dataset and get rid of all uninteresting data
birthorder = birthorder %>%
  filter(!is.na(pidlink)) %>% # no individuals who are only known from the pregnancy file
  filter(is.na(lifebirths) | lifebirths == 2) %>% # remove 7 and 2 individuals who are known as stillbirth or miscarriage but still have PID
  select(-lifebirths) %>% 
  filter(!is.na(mother_pidlink)) %>% 
  select(-father_pidlink) %>% 
  filter(is.na(any_multiple_birth) | any_multiple_birth != 1) %>% # remove families with twins/triplets/..
  select(-starts_with("age_"), -wave, -any_multiple_birth, -multiple_birth) %>% 
  mutate(money_spent_smoking_log = if_else(is.na(money_spent_smoking_log) & ever_smoked == 0, 0, money_spent_smoking_log),
         amount = if_else(is.na(amount) & ever_smoked == 0, 0, amount),
         amount_still_smokers = if_else(is.na(amount_still_smokers) &  still_smoking == 0, 0, amount_still_smokers),
         birthyear = lubridate::year(birthdate))

birthorder <- within(birthorder, {
  g_factor <- rowMeans(scale(cbind(raven_2015_old, math_2015_old, count_backwards,  words_delayed, adaptive_numbering)))
})

### Variables
birthorder = birthorder %>%
  mutate(
    # center variables that are used for analysis
  raven_2015_old_nonz = raven_2015_old,
  g_factor = scale(g_factor),
  raven_2015_old = scale(raven_2015_old),
  math_2015_old = scale(math_2015_old),
  count_backwards = scale(count_backwards),
  raven_2015_young = scale(raven_2015_young),
  math_2015_young = scale(math_2015_young),
  words_remembered_avg = scale(words_remembered_avg),
  words_immediate = scale(words_immediate),
  words_delayed = scale(words_delayed),
  adaptive_numbering = scale(adaptive_numbering),
  riskA = scale(riskA),
  riskB = scale(riskB),
  years_of_education_z = scale(years_of_education),
  Total_score_highest_z = scale(Total_score_highest),
  wage_last_month_z = scale(wage_last_month_log),
  wage_last_year_z = scale(wage_last_year_log),
  big5_ext = scale(big5_ext),
  big5_con = scale(big5_con),
  big5_agree = scale(big5_agree),
  big5_open = scale(big5_open),
  big5_neu = scale(big5_neu)
)
qplot(birthorder$male)
qplot(birthorder$g_factor)

### Birthorder and Sibling Count
birthorder = birthorder %>% 
  mutate(
# birthorder as factors with levels of 1, 2, 3, 4, 5, 5+
    birthorder_naive_factor = as.character(birthorder_naive),
    birthorder_naive_factor = ifelse(birthorder_naive > 5, "5+",
                                            birthorder_naive_factor),
    birthorder_naive_factor = factor(birthorder_naive_factor, 
                                            levels = c("1","2","3","4","5","5+")),
    sibling_count_naive_factor = as.character(sibling_count_naive),
    sibling_count_naive_factor = ifelse(sibling_count_naive > 5, "5+",
                                               sibling_count_naive_factor),
    sibling_count_naive_factor = factor(sibling_count_naive_factor, 
                                               levels = c("2","3","4","5","5+")),

    birthorder_uterus_alive_factor = as.character(birthorder_uterus_alive),
    birthorder_uterus_alive_factor = ifelse(birthorder_uterus_alive > 5, "5+",
                                            birthorder_uterus_alive_factor),
    birthorder_uterus_alive_factor = factor(birthorder_uterus_alive_factor, 
                                            levels = c("1","2","3","4","5","5+")),
    sibling_count_uterus_alive_factor = as.character(sibling_count_uterus_alive),
    sibling_count_uterus_alive_factor = ifelse(sibling_count_uterus_alive > 5, "5+",
                                               sibling_count_uterus_alive_factor),
    sibling_count_uterus_alive_factor = factor(sibling_count_uterus_alive_factor, 
                                               levels = c("2","3","4","5","5+")),
    birthorder_uterus_preg_factor = as.character(birthorder_uterus_preg),
    birthorder_uterus_preg_factor = ifelse(birthorder_uterus_preg > 5, "5+",
                                           birthorder_uterus_preg_factor),
    birthorder_uterus_preg_factor = factor(birthorder_uterus_preg_factor,
                                           levels = c("1","2","3","4","5","5+")),
    sibling_count_uterus_preg_factor = as.character(sibling_count_uterus_preg),
    sibling_count_uterus_preg_factor = ifelse(sibling_count_uterus_preg > 5, "5+",
                                              sibling_count_uterus_preg_factor),
    sibling_count_uterus_preg_factor = factor(sibling_count_uterus_preg_factor, 
                                              levels = c("2","3","4","5","5+")),
    birthorder_genes_factor = as.character(birthorder_genes),
    birthorder_genes_factor = ifelse(birthorder_genes >5 , "5+", birthorder_genes_factor),
    birthorder_genes_factor = factor(birthorder_genes_factor, 
                                     levels = c("1","2","3","4","5","5+")),
    sibling_count_genes_factor = as.character(sibling_count_genes),
    sibling_count_genes_factor = ifelse(sibling_count_genes >5 , "5+",
                                        sibling_count_genes_factor),
    sibling_count_genes_factor = factor(sibling_count_genes_factor, 
                                        levels = c("2","3","4","5","5+")),
    # interaction birthorder * siblingcout for each birthorder
    count_birthorder_naive =
      factor(str_replace(as.character(interaction(birthorder_naive_factor,                                                              sibling_count_naive_factor)),
                        "\\.", "/"),
                                           levels =   c("1/2","2/2", "1/3",  "2/3",
                                                        "3/3", "1/4", "2/4", "3/4", "4/4",
                                                        "1/5", "2/5", "3/5", "4/5", "5/5",
                                                        "1/5+", "2/5+", "3/5+", "4/5+",
                                                        "5/5+", "5+/5+")),
    count_birthorder_uterus_alive =
      factor(str_replace(as.character(interaction(birthorder_uterus_alive_factor,                                                              sibling_count_uterus_alive_factor)),
                        "\\.", "/"),
                                           levels =   c("1/2","2/2", "1/3",  "2/3",
                                                        "3/3", "1/4", "2/4", "3/4", "4/4",
                                                        "1/5", "2/5", "3/5", "4/5", "5/5",
                                                        "1/5+", "2/5+", "3/5+", "4/5+",
                                                        "5/5+", "5+/5+")),
    count_birthorder_uterus_preg =
      factor(str_replace(as.character(interaction(birthorder_uterus_preg_factor,                                                              sibling_count_uterus_preg_factor)), 
                         "\\.", "/"),
                                           levels =   c("1/2","2/2", "1/3",  "2/3",
                                                        "3/3", "1/4", "2/4", "3/4", "4/4",
                                                        "1/5", "2/5", "3/5", "4/5", "5/5",
                                                        "1/5+", "2/5+", "3/5+", "4/5+",
                                                        "5/5+", "5+/5+")),
    count_birthorder_genes =
      factor(str_replace(as.character(interaction(birthorder_genes_factor,                                                              sibling_count_genes_factor)), "\\.", "/"),
                                           levels =   c("1/2","2/2", "1/3",  "2/3",
                                                        "3/3", "1/4", "2/4", "3/4", "4/4",
                                                        "1/5", "2/5", "3/5", "4/5", "5/5",
                                                        "1/5+", "2/5+", "3/5+", "4/5+",
                                                        "5/5+", "5+/5+")))

birthorder <- birthorder %>%
                     mutate(sibling_count = sibling_count_naive_factor,
                            birth_order_nonlinear = birthorder_naive_factor,
                            birth_order = birthorder_naive,
                            count_birth_order = count_birthorder_naive)

```

## How many siblings with data do we retain in each family?
```{r}
crosstabs(~is.na(birthorder$birthorder_naive) + is.na(birthorder$g_factor))
crosstabs(~is.na(birthorder$birthorder_uterus_preg) + is.na(birthorder$g_factor))

birthorder %>% filter(!is.na(birthorder_naive), !is.na(g_factor)) %>% group_by(mother_pidlink) %>% summarise(with_data = n(), all = mean(sibling_count_uterus_preg)) -> counts
birthorder %>% filter(!is.na(birthorder_naive)) %>% group_by(mother_pidlink) %>% summarise(with_data = n(), all = mean(sibling_count_uterus_preg)) -> counts1
```

In our test sample families with an average size of `r mean(counts$all,na.rm= T)` siblings, we retain `r mean(counts$with_data)`.

In the original sample families with an average size of `r mean(counts1$all,na.rm= T)` siblings, we retain `r mean(counts1$with_data)`.

```{r}
ggplot(counts, aes(all, with_data)) + geom_jitter(alpha = 0.1) + geom_smooth() + scale_x_continuous(breaks=1:15) + scale_y_continuous(breaks=1:15)
```


## Intelligence {.active .tabset}
### g-factor {.tabset}

```{r}
birthorder <- birthorder %>% mutate(outcome = g_factor)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```


### Raven {.tabset}

```{r}
birthorder <- birthorder %>% mutate(outcome = raven_2015_old)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```

### Numeracy {.tabset}

```{r}
birthorder <- birthorder %>% mutate(outcome = math_2015_old)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```


### Counting backwards {.tabset}

```{r}
birthorder <- birthorder %>% mutate(outcome = count_backwards)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```

### Immediate word recall {.tabset}

```{r}
birthorder <- birthorder %>% mutate(outcome = words_immediate)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```

### Delayed word recall {.tabset}
```{r}
birthorder <- birthorder %>% mutate(outcome = words_delayed)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```

### Adaptive Numbering {.tabset}

```{r}
birthorder <- birthorder %>% mutate(outcome = adaptive_numbering)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```

## Personality {.tabset}
### Extraversion {.tabset}


```{r}
birthorder <- birthorder %>% mutate(outcome = big5_ext)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```


### Neuroticism {.tabset}


```{r}
birthorder <- birthorder %>% mutate(outcome = big5_neu)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```


### Conscientiousness {.tabset}


```{r}
birthorder <- birthorder %>% mutate(outcome = big5_con)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```


### Agreeableness {.tabset}


```{r}
birthorder <- birthorder %>% mutate(outcome = big5_agree)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```


### Openness {.tabset}


```{r}
birthorder <- birthorder %>% mutate(outcome = big5_open)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```


## Risk preference {.tabset}

### Risk A {.tabset}


```{r}
birthorder <- birthorder %>% mutate(outcome = riskA)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```


### Risk B {.tabset}


```{r}
birthorder <- birthorder %>% mutate(outcome = riskB)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```


## Educational Attainment {.tabset}
### Years of Education - z-standardized {.tabset}


```{r}
birthorder <- birthorder %>% mutate(outcome = years_of_education_z)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```

#### Years of education - non-standardized {.tabset}

```{r}
birthorder <- birthorder %>% mutate(outcome = years_of_education)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```

### EBTANAS {.tabset}

```{r}
birthorder <- birthorder %>% mutate(outcome = Total_score_elemenatry)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```


## Income last month {.tabset}

### z-standardized {.tabset}

```{r}
birthorder <- birthorder %>% mutate(outcome = wage_last_month_z)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```

## Income last year {.tabset}

### z-standardized {.tabset}

```{r}
birthorder <- birthorder %>% mutate(outcome = wage_last_year_z)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```



## Work {.tabset}
### Self-Employment - non standardized {.tabset}


```{r}
birthorder <- birthorder %>% mutate(outcome = Self_employed)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```


### Child Labour {.tabset}
```{r}
birthorder <- birthorder %>% mutate(outcome = total_worked)
model = lmer(outcome ~ birth_order + age + male + sibling_count + (1 | mother_pidlink),
                   data = birthorder)
compare_birthorder_specs(model)
```

