---
title: "4. Analyses Imputed Data"
author: "Laura Botzet & Ruben Arslan"
output: html_document
editor_options: 
  chunk_output_type: console
---

# <span style="color:#FFD92F">Birth Order Effects Imputed Data</span> {.tabset}

## Helper

```{r helper}
source("0_helpers.R")
opts_chunk$set(warning = FALSE, error = T)
detach("package:lmerTest", unload = TRUE)
```


## Functions

```{r}
options(mc.cores=4)
n_imputations.mitml = function(imps) {
  imps$iter$m
}
n_imputations.mids = function(imps) {
  imps$m
}
n_imputations.amelia = function(imps) {
  length(imps$imputations)
}
n_imputations = function(imps) { UseMethod("n_imputations") }

complete.mitml = mitml::mitmlComplete
complete.mids = mice::complete
complete = function(...) { UseMethod("complete") }

library(mice)
library(miceadds)
library(mitml)
```


## Load data

```{r Load Data}
birthorder_missing = readRDS("data/alldata_birthorder_missing.rds")
birthorder = readRDS("data/alldata_birthorder_i_ml5.rds")
```

## Data preparations

```{r data preparations}
birthorder <- mids2mitml.list(birthorder)

## Calculate g_factors (for now with rowMeans: better to run factor analysis here as well?)
birthorder <- within(birthorder, {
  g_factor_2015_old <- rowMeans(scale(cbind(raven_2015_old, math_2015_old, count_backwards,  words_delayed, adaptive_numbering)))
})

birthorder <- within(birthorder, {
  g_factor_2015_young <- rowMeans(scale(cbind(raven_2015_young, math_2015_young)))
})

birthorder <- within(birthorder, {
  g_factor_2007_old <- rowMeans(scale(cbind(raven_2007_old, math_2007_old)))
})

birthorder <- within(birthorder, {
  g_factor_2007_young <- rowMeans(scale(cbind(raven_2007_young, math_2007_young)))
})


birthorder = birthorder %>% tbl_df.mitml.list

### Birthorder and Sibling Count
birthorder_data = birthorder[[1]]

birthorder = birthorder %>%
  mutate(birthorder_genes = ifelse(birthorder_genes < 1, 1, birthorder_genes),
         birthorder_naive = ifelse(birthorder_naive < 1, 1, birthorder_naive),
         birthorder_genes = round(birthorder_genes),
         birthorder_naive = round(birthorder_naive),
         sibling_count_genes = ifelse(sibling_count_genes < 2, 2, sibling_count_genes),
         sibling_count_naive = ifelse(sibling_count_naive < 2, 2, sibling_count_naive),
         sibling_count_genes = round(sibling_count_genes),
         sibling_count_naive = round(sibling_count_naive)
         ) %>% 
  group_by(mother_pidlink) %>% 
  # make sure sibling count is consistent with birth order (for which imputation model seems to work better)
  mutate(sibling_count_genes = max(birthorder_genes, sibling_count_genes)) %>% 
  ungroup()


birthorder = birthorder %>% 
  mutate(
# birthorder as factors with levels of 1, 2, 3, 4, 5, 5+
    birthorder_naive_factor = as.character(birthorder_naive),
    birthorder_naive_factor = ifelse(birthorder_naive > 5, "5+",
                                            birthorder_naive_factor),
    birthorder_naive_factor = factor(birthorder_naive_factor, 
                                            levels = c("1","2","3","4","5","5+")),
    sibling_count_naive_factor = as.character(sibling_count_naive),
    sibling_count_naive_factor = ifelse(sibling_count_naive > 5, "5+",
                                               sibling_count_naive_factor),
    sibling_count_naive_factor = factor(sibling_count_naive_factor, 
                                               levels = c("2","3","4","5","5+")),
    birthorder_genes_factor = as.character(birthorder_genes),
    birthorder_genes_factor = ifelse(birthorder_genes >5 , "5+", birthorder_genes_factor),
    birthorder_genes_factor = factor(birthorder_genes_factor, 
                                     levels = c("1","2","3","4","5","5+")),
    sibling_count_genes_factor = as.character(sibling_count_genes),
    sibling_count_genes_factor = ifelse(sibling_count_genes >5 , "5+",
                                        sibling_count_genes_factor),
    sibling_count_genes_factor = factor(sibling_count_genes_factor, 
                                        levels = c("2","3","4","5","5+")),
    # interaction birthorder * siblingcout for each birthorder
    count_birthorder_naive =
      factor(str_replace(as.character(interaction(birthorder_naive_factor,                                                              sibling_count_naive_factor)),
                        "\\.", "/"),
                                           levels =   c("1/2","2/2", "1/3",  "2/3",
                                                        "3/3", "1/4", "2/4", "3/4", "4/4",
                                                        "1/5", "2/5", "3/5", "4/5", "5/5",
                                                        "1/5+", "2/5+", "3/5+", "4/5+",
                                                        "5/5+", "5+/5+")),
    count_birthorder_genes =
      factor(str_replace(as.character(interaction(birthorder_genes_factor,                                                              sibling_count_genes_factor)), "\\.", "/"),
                                           levels =   c("1/2","2/2", "1/3",  "2/3",
                                                        "3/3", "1/4", "2/4", "3/4", "4/4",
                                                        "1/5", "2/5", "3/5", "4/5", "5/5",
                                                        "1/5+", "2/5+", "3/5+", "4/5+",
                                                        "5/5+", "5+/5+")))



### Variables
birthorder = birthorder %>%
  mutate(
    g_factor_2015_old = (g_factor_2015_old - mean(g_factor_2015_old, na.rm=T)) / sd(g_factor_2015_old, na.rm=T),
    g_factor_2015_young = (g_factor_2015_young - mean(g_factor_2015_young, na.rm=T)) / sd(g_factor_2015_young, na.rm=T),
    g_factor_2007_old = (g_factor_2007_old - mean(g_factor_2007_old, na.rm=T)) / sd(g_factor_2007_old, na.rm=T),
g_factor_2007_young = (g_factor_2007_young - mean(g_factor_2007_young, na.rm=T)) / sd(g_factor_2007_young, na.rm=T),
    raven_2015_old = (raven_2015_old - mean(raven_2015_old, na.rm=T)) / sd(raven_2015_old),
    math_2015_old = (math_2015_old - mean(math_2015_old, na.rm=T)) / sd(math_2015_old, na.rm=T),
    raven_2015_young = (raven_2015_young - mean(raven_2015_young, na.rm=T)) / sd(raven_2015_young),
    math_2015_young = (math_2015_young - mean(math_2015_young, na.rm=T)) /
      sd(math_2015_young, na.rm=T),
    raven_2007_old = (raven_2007_old - mean(raven_2007_old, na.rm=T)) / sd(raven_2007_old),
math_2007_old = (math_2007_old - mean(math_2007_old, na.rm=T)) / sd(math_2007_old, na.rm=T),
raven_2007_young = (raven_2007_young - mean(raven_2007_young, na.rm=T)) / sd(raven_2007_young),
math_2007_young = (math_2007_young - mean(math_2007_young, na.rm=T)) /
  sd(math_2007_young, na.rm=T),
    count_backwards = (count_backwards - mean(count_backwards, na.rm=T)) /
      sd(count_backwards, na.rm=T),
    words_delayed = (words_delayed - mean(words_delayed, na.rm=T)) /
      sd(words_delayed, na.rm=T),
    adaptive_numbering = (adaptive_numbering - mean(adaptive_numbering, na.rm=T)) /
      sd(adaptive_numbering, na.rm=T),
    big5_ext = (big5_ext - mean(big5_ext, na.rm=T)) / sd(big5_ext, na.rm=T),
    big5_con = (big5_con - mean(big5_con, na.rm=T)) / sd(big5_con, na.rm=T),
    big5_agree = (big5_agree - mean(big5_agree, na.rm=T)) / sd(big5_agree, na.rm=T),
    big5_open = (big5_open - mean(big5_open, na.rm=T)) / sd(big5_open, na.rm=T),
    big5_neu = (big5_neu - mean(big5_neu, na.rm=T)) / sd(big5_neu, na.rm=T),
    riskA = (riskA - mean(riskA, na.rm=T)) / sd(riskA, na.rm=T),
    riskB = (riskB - mean(riskB, na.rm=T)) / sd(riskB, na.rm=T),
    years_of_education_z = (years_of_education - mean(years_of_education, na.rm=T)) /
      sd(years_of_education, na.rm=T)
)

        
data_used <- birthorder %>%
  mutate(sibling_count = sibling_count_genes_factor,
         birth_order_nonlinear = birthorder_genes_factor,
         birth_order = birthorder_genes,
         count_birth_order = count_birthorder_genes)

```

## Intelligence {.active .tabset}
### g-factor 2015 old {.tabset}

```{r gfold15o}
# Specify your outcome
data_used = data_used %>% mutate(outcome = g_factor_2015_old)
data_used_plots <- data_used[[1]]

compare_birthorder_imputed(data_used = data_used)

```

### g-factor 2015 young {.tabset}

```{r gf15y}
# Specify your outcome
data_used = data_used %>% mutate(outcome = g_factor_2015_young)
data_used_plots <- data_used[[1]]

compare_birthorder_imputed(data_used = data_used)

```

## Personality {.tabset}

### Extraversion {.tabset}

```{r extra}
# Specify your outcome
data_used = data_used %>% mutate(outcome = big5_ext)
data_used_plots <- data_used[[1]]

compare_birthorder_imputed(data_used = data_used)

```


### Neuroticism {.tabset}

```{r neuro}
# Specify your outcome
data_used = data_used %>% mutate(outcome = big5_neu)
compare_birthorder_imputed(data_used = data_used)

```


### Conscientiousness {.tabset}

```{r consc}
# Specify your outcome
data_used = data_used %>% mutate(outcome = big5_con)
data_used_plots <- data_used[[1]]

compare_birthorder_imputed(data_used = data_used)

```


### Agreeableness {.tabset}

```{r agree}
# Specify your outcome
data_used = data_used %>% mutate(outcome = big5_agree)
data_used_plots <- data_used[[1]]

compare_birthorder_imputed(data_used = data_used)

```


### Openness {.tabset}

```{r open}
# Specify your outcome
data_used = data_used %>% mutate(outcome = big5_open)
data_used_plots <- data_used[[1]]

compare_birthorder_imputed(data_used = data_used)

```

## Riskpreference {.tabset}

### Risk A {.tabset}

```{r riska}
# Specify your outcome
data_used = data_used %>% mutate(outcome = riskA)
data_used_plots <- data_used[[1]]

compare_birthorder_imputed(data_used = data_used)

```

### Risk B {.tabset}

```{r riskb}
# Specify your outcome
data_used = data_used %>% mutate(outcome = riskB)
data_used_plots <- data_used[[1]]

compare_birthorder_imputed(data_used = data_used)
```

## Educational Attainment {.tabset}
### Years of Education - z-standardized{.tabset}

```{r edu}
# Specify your outcome
data_used = data_used %>% mutate(outcome = years_of_education_z)
data_used_plots <- data_used[[1]]

compare_birthorder_imputed(data_used = data_used)

```


