---
output: html_document
editor_options: 
  chunk_output_type: console
---
# <span style="color:#A6D854">Birth Order Effects</span> {.tabset}

## Helper
```{r helper}
source("0_helpers.R")
```


## Functions

```{r}

n_imputations.mitml = function(imps) {
  imps$iter$m
}
n_imputations.mids = function(imps) {
  imps$m
}
n_imputations.amelia = function(imps) {
  length(imps$imputations)
}
n_imputations = function(imps) { UseMethod("n_imputations") }

complete.mitml = mitml::mitmlComplete
complete.mids = mice::complete
complete = function(...) { UseMethod("complete") }

library(mice)
library(miceadds)
library(mitml)
```


## Load data
```{r Load Data}
birthorder = readRDS("data/alldata_birthorder_i.rds")
```

## Data preparations
```{r data preparations}
birthorder <- mids2mitml.list(birthorder)
birthorder <- within(birthorder, {
  g_factor_2015 <- rowMeans(scale(cbind(raven_2015_old, math_2015_old, count_backwards,  words_delayed, adaptive_numbering)))
})
birthorder = birthorder %>% tbl_df.mitml.list

### Birthorder and Sibling Count
birthorder = birthorder %>% 
  mutate(
# birthorder as factors with levels of 1, 2, 3, 4, 5, 5+
    birthorder_naive_factor = as.character(birthorder_naive),
    birthorder_naive_factor = ifelse(birthorder_naive > 5, "5+",
                                            birthorder_naive_factor),
    birthorder_naive_factor = factor(birthorder_naive_factor, 
                                            levels = c("1","2","3","4","5","5+")),
    sibling_count_naive_factor = as.character(sibling_count_naive),
    sibling_count_naive_factor = ifelse(sibling_count_naive > 5, "5+",
                                               sibling_count_naive_factor),
    sibling_count_naive_factor = factor(sibling_count_naive_factor, 
                                               levels = c("2","3","4","5","5+")),
    birthorder_genes_factor = as.character(birthorder_genes),
    birthorder_genes_factor = ifelse(birthorder_genes >5 , "5+", birthorder_genes_factor),
    birthorder_genes_factor = factor(birthorder_genes_factor, 
                                     levels = c("1","2","3","4","5","5+")),
    sibling_count_genes_factor = as.character(sibling_count_genes),
    sibling_count_genes_factor = ifelse(sibling_count_genes >5 , "5+",
                                        sibling_count_genes_factor),
    sibling_count_genes_factor = factor(sibling_count_genes_factor, 
                                        levels = c("2","3","4","5","5+")),
    # interaction birthorder * siblingcout for each birthorder
    count_birthorder_naive =
      factor(str_replace(as.character(interaction(birthorder_naive_factor,                                                              sibling_count_naive_factor)),
                        "\\.", "/"),
                                           levels =   c("1/2","2/2", "1/3",  "2/3",
                                                        "3/3", "1/4", "2/4", "3/4", "4/4",
                                                        "1/5", "2/5", "3/5", "4/5", "5/5",
                                                        "1/5+", "2/5+", "3/5+", "4/5+",
                                                        "5/5+", "5+/5+")),
    count_birthorder_genes =
      factor(str_replace(as.character(interaction(birthorder_genes_factor,                                                              sibling_count_genes_factor)), "\\.", "/"),
                                           levels =   c("1/2","2/2", "1/3",  "2/3",
                                                        "3/3", "1/4", "2/4", "3/4", "4/4",
                                                        "1/5", "2/5", "3/5", "4/5", "5/5",
                                                        "1/5+", "2/5+", "3/5+", "4/5+",
                                                        "5/5+", "5+/5+")))



### Variables
birthorder = birthorder %>%
  mutate(
    g_factor_2015 = (g_factor_2015 - mean(g_factor_2015, na.rm=T)) / sd(g_factor_2015, na.rm=T),
    raven_2015_old = (raven_2015_old - mean(raven_2015_old, na.rm=T)) / sd(raven_2015_old),
    math_2015_old = (math_2015_old - mean(math_2015_old, na.rm=T)) / sd(math_2015_old, na.rm=T),
    raven_2015_young = (raven_2015_young - mean(raven_2015_young, na.rm=T)) / sd(raven_2015_young),
    math_2015_young = (math_2015_young - mean(math_2015_young, na.rm=T)) /
      sd(math_2015_young, na.rm=T),
    count_backwards = (count_backwards - mean(count_backwards, na.rm=T)) /
      sd(count_backwards, na.rm=T),
    words_delayed = (words_delayed - mean(words_delayed, na.rm=T)) /
      sd(words_delayed, na.rm=T),
    adaptive_numbering = (adaptive_numbering - mean(adaptive_numbering, na.rm=T)) /
      sd(adaptive_numbering, na.rm=T),
    big5_ext = (big5_ext - mean(big5_ext, na.rm=T)) / sd(big5_ext, na.rm=T),
    big5_con = (big5_con - mean(big5_con, na.rm=T)) / sd(big5_con, na.rm=T),
    big5_agree = (big5_agree - mean(big5_agree, na.rm=T)) / sd(big5_agree, na.rm=T),
    big5_open = (big5_open - mean(big5_open, na.rm=T)) / sd(big5_open, na.rm=T),
    big5_neu = (big5_neu - mean(big5_neu, na.rm=T)) / sd(big5_neu, na.rm=T),
    riskA = (riskA - mean(riskA, na.rm=T)) / sd(riskA, na.rm=T),
    riskB = (riskB - mean(riskB, na.rm=T)) / sd(riskB, na.rm=T),
    years_of_education_z = (years_of_education - mean(years_of_education, na.rm=T)) /
      sd(years_of_education, na.rm=T),
    wage_last_month_z = (wage_last_month_log - mean(wage_last_month_log, na.rm=T)) /
      sd(wage_last_month_log, na.rm=T),
    wage_last_year_z = (wage_last_year_log - mean(wage_last_year_log, na.rm=T)) /
      sd(wage_last_year_log, na.rm=T)
        )

```

## Intelligence {.active .tabset}
### g-factor {.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = g_factor_2015)
compare_birthorder_imputed(data_used = birthorder)

```

### Raven {.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = raven_2015_old)
compare_birthorder_imputed(data_used = birthorder)

```

### Numeracy {.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = math_2015_old)
compare_birthorder_imputed(data_used = birthorder)

```

### Counting Backwards {.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = count_backwards)
compare_birthorder_imputed(data_used = birthorder)

```

### Delayed Word Recall {.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = words_delayed)
compare_birthorder_imputed(data_used = birthorder)

```


### Adaptive Numbering {.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = adaptive_numbering)
compare_birthorder_imputed(data_used = birthorder)

```

## Personality {.tabset}

### Extraversion {.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = big5_ext)
compare_birthorder_imputed(data_used = birthorder)

```


### Neuroticism {.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = big5_neu)
compare_birthorder_imputed(data_used = birthorder)

```


### Conscientiousness {.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = big5_neu)
compare_birthorder_imputed(data_used = birthorder)

```


### Agreeableness {.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = big5_agree)
compare_birthorder_imputed(data_used = birthorder)

```


### Openness {.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = Openness)
compare_birthorder_imputed(data_used = birthorder)

```

## Riskpreference {.tabset}
### Risk A {.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = riskA)
compare_birthorder_imputed(data_used = birthorder)

```

### Risk B {.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = riskB)
compare_birthorder_imputed(data_used = birthorder)

```

## Educational Attainment {.tabset}
### Years of Education - z-standardized{.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = years_of_education_z)
compare_birthorder_imputed(data_used = birthorder)

```
### Years of Education - non z-standardized{.tabset}

```{r}
# Specify your outcome
birthorder = birthorder %>% mutate(outcome = years_of_education)
compare_birthorder_imputed(data_used = birthorder)

```
