---
title: "Birth order effects in the IFLS"
bibliography: packrat_bibliography.bibtex
csl: apa-custom-no-issue.csl
nocite: |
    @R @packrat @rmarkdown @knitr @formr @pander @lubridate @stringr @broom @effects @ggplot2 @tidyr @dplyr @formatR @devtools @haven @GPArotation @psych @lavaan @Hmisc @plyr @apaTables @lme4 @lmerTest @forcats @cowplot @gridExtra
---

# Indonesian Family Life Study {.tabset .tabset-sticky}

More information about the [IFLS](http://www.rand.org/labor/FLS/IFLS.html).

```{r message=FALSE, warning=FALSE}
source("0_helpers.R")
```

We used these [helper functions](0_helpers.html) and 
this [knitr component](_test_outcome.html). These are the [.Rprofile settings](.Rprofile.html)

1. First we [wrangled the data](1_data_import.html).
2. Then we looked at some [descriptives](2_demographics.html).
3. Then we [analysed](4_analyses.html) it.

## Authors & Acknowledgements {.tabset .tabset-sticky}

### Authors
Laura J. Botzet & Ruben C. Arslan.

This supplementary website has been archived on Zenodo.org [![DOI]()]().

### Acknowledgements {.active}


#### Package bibliography {.tabset .tabset-sticky}

##### Session info
```{r warning=FALSE,message=FALSE}
# Make packrat bibliography
packrat_bibliography(overwrite_bib = TRUE, silent = TRUE)
# Turn the .R files into .Rmd files, turn those into .html, remove the .Rmd files
spin_R_files_to_site_html()
sessionInfo()
```

##### References {.active}
